---
title: "Prevalence of Diabetes in the United States"
author: Bobby Decker
subtitle: Visualizing the spread of diabetes in the United States from 2010-2018
---

# Introduction


Diabetes has been spreading rapidly throughout the United States through the last 20 years. Type 2 Diabetes is associated with obesity which stems from poor eating habits and lack of consistent physical activity. It has been studied that more than 100 million U.S. adults are living with diabetes or are considered prediabetic. 

As of 2015, 30.3 million Americans - 9.4 percent of the U.S. population - are diabetic. Diabetes is a serious disease that can be mitigated through increased physical activity to maintain safe and consistent blood sugar levels. 

# Materials and methods


For the purpose of this study, a few different visualizations of diabetes and levels of physical activity will be created through the R packages such as ggplot2. An animated map showing the change in spatial distribution of diabetes will be produced using the R package gganimate. 

Code: The code associated with the project is well organized and easy to follow. Demonstrates mastery of R graphics and functions.

Data: The underlying data are publicly accessible via the CDC GIS website and can also be downloaded/accessed within the Rmd script.

First we must load the necessary R packages
```{r}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(janitor)
library(gganimate)
```

See [http://rmarkdown.rstudio.com/](http://rmarkdown.rstudio.com/) for all the amazing things you can do.


Here's my first code chunk.
```{r}

```

Load any required packages in a code chunk (you may need to install some packages):

```{r, message=F, warning=F}
library(tidyverse)
library(leaflet)
library(kableExtra)
knitr::opts_chunk$set(cache=TRUE)  # cache the results for quick compiling
```

## Download and clean all required data
```{r}
n=20
data=data.frame(x=runif(n,-180,180),
                y=runif(n,-60,60),
                size = runif(n, 5, 20),
                category = factor(
                  sample(letters[1:5], n, replace = TRUE)
                  ),
                value = rnorm(n))
```

```{r, results='asis'}
data %>% 
  slice(1:10) %>% #show only 1:n rows
  kable(digits=2,align="c")%>% #make table and round to two digits
  kable_styling(bootstrap_options = 
                  c("striped", "hover", "condensed", "responsive")) #apply other formatting
```

Add any additional processing steps here.

# Results

[~200 words]

Tables and figures (maps and other graphics) are carefully planned to convey the results of your analysis. Intense exploration and evidence of many trials and failures. The author looked at the data in many different ways before coming to the final presentation of the data.

Show tables, plots, etc. and describe them.

```{r, fig.width=6, fig.height=3, fig.cap="Map of completely random data"}
m <- leaflet(data) %>% 
  addTiles() %>% 
  addCircleMarkers(~x, ~y, radius = ~size,color = ~as.factor(category)) %>% 
  addPopups(~x[2], ~y[2], "Random popup")
m  # a map with the default OSM tile layer
```


```{r}
data %>% 
  ggplot(aes(x=x,y=y,col=category))+
  geom_point()
```

# Conclusions

[~200 words]

Clear summary adequately describing the results and putting them in context. Discussion of further questions and ways to continue investigation.

# References

All sources are cited in a consistent manner
