---
title: "Prevalence of Diabetes in the United States"
author: Bobby Decker
subtitle: Visualizing the spread of diabetes in the United States from 2010-2018
---

# Introduction


Diabetes has been spreading rapidly throughout the United States through the last 20 years. Type 2 Diabetes is associated with obesity which stems from poor eating habits and lack of consistent physical activity. It has been studied that more than 100 million U.S. adults are living with diabetes or are considered prediabetic. 

As of 2015, 30.3 million Americans - 9.4 percent of the U.S. population - are diabetic. Diabetes is a serious disease that can be mitigated through increased physical activity to maintain safe and consistent blood sugar levels. 

For the purpose of this project I am going to create visualizations that compare the levels of diagnosed diabetes percentage with the percentage levels of physical inactivity in the United States and in NY State. It is hypothesized that we will see an overall increase in levels of Diagnosed Diabetes percentage. Finally, it is also hypothesized that we will observe a correlation between diagnosed diabetes percentage and levels of physical inactivity.

# Materials and methods


For the purpose of this study, a few different visualizations of diabetes and levels of physical activity will be created through the R packages such as ggplot2. First I will compare the overall diabetes diagnosed percentages throughout the United States from 2010-2018. Then I will produce a visualization of the diagnosed diabetes percentages in NY in 2018 and compare it with the percentage levels of physical inactivity collected in the same year. 

First we must load the necessary R packages
```{r, message=FALSE, warning=FALSE, results=FALSE}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(janitor)
library(usmap)
library(ggthemes)
library(maps)
library(patchwork)
```


Next we must read and transform the diabetes and physical inactivity data from the CDC's website https://gis.cdc.gov/grasp/diabetes/diabetesatlas-sdoh.html
```{r, message=FALSE, warning=FALSE} 
diabetes_2010<- read.csv("DiabetesAtlasData_2010.csv")
diabetes_2018 <- read.csv("DiabetesAtlasData_2018.csv")%>%
    rename(fips=County_FIPS_2018)

inactivity_2010 <- read.csv("DiabetesAtlasData_activity_2010.csv")
inactivity_2018 <- read.csv("DiabetesAtlasData_activity_2018.csv")%>%
    rename(fips=County_FIPS)


diabetes_data<-cbind(diabetes_2010,diabetes_2018)%>%
  select(County_FIPS, Diagnosed.Diabetes.Percentage,Diagnosed.Diabetes.Percentage_2018)%>%
  rename(fips=County_FIPS)
inactivity_data<-cbind(inactivity_2010,inactivity_2018)
```


```{r, results='asis'}

```


# Results


First we are going to produce a map of the distribution of Diabetes in the United States in 2010
```{r, message=FALSE, warning=FALSE}
states <- plot_usmap("states", 
                     color = "red",
                     fill = alpha(0.01)) #this parameter is necessary to get counties to show on top of states
counties <- plot_usmap(data = diabetes_data, 
                       values = "Diagnosed.Diabetes.Percentage",
                       color = "black",
                       size = 0.1)

diabetes<-ggplot() +  
  geom_polygon(data=counties[[1]], 
               aes(x=x, 
                   y=y, 
                   group=group, 
                   fill = counties[[1]]$Diagnosed.Diabetes.Percentage), 
               color = "black",
               size = 0.1) +  
  geom_polygon(data=states[[1]], 
               aes(x=x, 
                   y=y, 
                   group=group), 
               color = "red", 
               fill = alpha(0.01)) + 
  coord_equal() +
  theme_map() +
  theme(legend.position="bottom") +
  scale_fill_gradient(low='white', high='grey20')+
  labs(
    title="US Diabetes Distribution 2010"
  )

diabetes
```


Now for comparison we are going to produce a similar map of the distribution of Diabetes in the United States in 2018
```{r, message=FALSE, warning=FALSE}
states <- plot_usmap("states", 
                     color = "red",
                     fill = alpha(0.01)) #this parameter is necessary to get counties to show on top of states
counties1 <- plot_usmap(data = diabetes_data, 
                       values = "Diagnosed.Diabetes.Percentage_2018",
                       color = "black",
                       size = 0.1)

diabetes1<-ggplot() +  
  geom_polygon(data=counties[[1]], 
               aes(x=x, 
                   y=y, 
                   group=group, 
                   fill = counties[[1]]$Diagnosed.Diabetes.Percentage_2018), 
               color = "black",
               size = 0.1) +  
  geom_polygon(data=states[[1]], 
               aes(x=x, 
                   y=y, 
                   group=group), 
               color = "red", 
               fill = alpha(0.01)) + 
  coord_equal() +
  theme_map() +
  theme(legend.position="bottom") +
  scale_fill_gradient(low='white', high='grey20')+
  labs(
    title="US Diabetes Distribution 2018"
  )

diabetes1
```

Now we are going use the patchwork package to compare these two maps side by side.
```{r}
diabetes+diabetes1
```

Next, we are going to create a county map of NYS showing diagnosed diabetes percentages in 2018
```{r,message=FALSE, warning=FALSE}
NY_diabetes<-plot_usmap(data = diabetes_2018, values = "Diagnosed.Diabetes.Percentage_2018",include = c("NY"), color = "red") +
   scale_fill_continuous(
    low = "white", high = "red", name = "Diagnosed Diabetes Percentage", label = scales::comma)+
  labs(title = "Diagnosed Diabetes Population In NY 2018",)+
    theme(legend.position = "bottom")
NY_diabetes
```

We are going to produce a similar comparison map, this time shoing the percentages of physical inactivity levels in counties of NYS in 2018
```{r,message=FALSE, warning=FALSE}
NY_Inactivity<-plot_usmap(data = inactivity_2018, values = "Physical.Inactivity.Percentage",include = c("NY"), color = "red") +
   scale_fill_continuous(
    low = "white", high = "red", name = "Physcial Inactivity Perecentage", label = scales::comma)+
  labs(title = "Physcial Inactivity Percentage In NY 2018",)+
    theme(legend.position = "bottom")
NY_diabetes
```
Here are the two maps together for comparison.
```{r,message=FALSE, warning=FALSE}
NY_diabetes+NY_Inactivity+
      theme(legend.position = "right")

```
# Conclusions
Through the visualizations created we can see there has been a steady observable increase in diabetes diagnosis percentage in many counties in the US from 2010-2018.Also, through our visualization of NYS we can see that there is a correlation between physical inactivity and diagnosed diabetes percentage. These results are consistent with the data and other sources researching diabetes and related health effects. 

# References

“Social Determinants of Health - United States Diabetes Surveillance System.” Centers for Disease Control and Prevention, Centers for Disease Control and Prevention, https://gis.cdc.gov/grasp/diabetes/diabetesatlas-sdoh.html. 